
---
title: "06 Temperature-Infection Interaction Analysis"
author: "Tegan Jarvis"
output: html_document
---

```{r}
library(DESeq2)
library(EnhancedVolcano)
```

```{r}
counts <- readRDS("results/counts_matrix.rds")
metadata <- readRDS("results/metadata.rds")
```

```{r}
dds_interaction <- DESeqDataSetFromMatrix(
  countData = counts,
  colData = metadata,
  design = ~ temperature * infection
)
```

```{r}
dds_interaction <- DESeq(dds_interaction)
```

```{r}
resultsNames(dds_interaction)
```

```{r}
res_interaction <- results(
  dds_interaction,
  name = "temperature5.infectionNI"
)

summary(res_interaction)
```
```{r}
res_interaction_df <- as.data.frame(res_interaction)
res_interaction_df$ensembl_gene_id <- rownames(res_interaction_df)

sig_interaction <- subset(
  res_interaction_df,
  !is.na(padj) &
    padj < 0.05 &
    abs(log2FoldChange) > 1
)

nrow(sig_interaction)
```


```{r}
head(res_interaction[order(res_interaction$padj), ])
```

```{r}
png("volcano_ana.png",
    width = 1200,
    height = 1000,
    res = 150)

EnhancedVolcano(
  res_interaction_df,

  lab = rownames(res_interaction_df),

  x = "log2FoldChange",
  y = "padj",

  pCutoff = 0.05,
  FCcutoff = 1,

  title = "Temperature-Infection Interaction",

  subtitle = "Genes whose infection response changes with temperature"
)

dev.off()

```

far right: stronger infection response at 5°C
far left: weaker or reversed response at 5°C

```{r}
dir.create("results", showWarnings = FALSE)

saveRDS(dds_interaction, "results/dds_interaction.rds")
saveRDS(res_interaction, "results/res_interaction.rds")
saveRDS(sig_interaction, "results/sig_interaction_genes.rds")

write.csv(
  res_interaction_df,
  "results/interaction_results.csv",
  row.names = FALSE
)

write.csv(
  sig_interaction,
  "results/significant_interaction_genes.csv",
  row.names = FALSE
)
```

The interaction analysis tested whether the transcriptional response to infection differed between 37°C and 5°C. Compared with the broad infection effect, fewer genes showed significant interaction effects. This is suggesting that temperature modifies a subset of infection-responsive genes rather than globally changing the entire infection response.







