---
title: "07 Replication of Published Figures"
author: "Tegan Jarvis"
output: html_document
---

This notebook is to create the original paper’s PCA + sample distance heatmap as closely as possible, then compare results.

Harazim M, Perrot J, Varet H, Bourhy H, Lannoy J, Pikula J, Seidlová V, Dacheux L, Martínková N. Transcriptomic responses of bat cells to European bat lyssavirus 1 infection under conditions simulating euthermia and hibernation. BMC Immunol. 2023 Apr 21;24(1):7. doi: 10.1186/s12865-023-00542-7. PMID: 37085747; PMCID: PMC10120247.

```{r}
library(DESeq2)
library(ggplot2)
library(pheatmap)
library(RColorBrewer)
library(patchwork)
```

```{r}
vsd <- readRDS("results/vsd.rds")
metadata <- readRDS("results/metadata.rds")
```

```{r}
metadata$design <- paste0(
  metadata$infection,
  "_t",
  metadata$temperature
)

head(metadata)
```


```{r}
colData(vsd)$design <- paste0(
  colData(vsd)$infection,
  "_t",
  colData(vsd)$temperature
)

colData(vsd)$replicate <- rep(c("r1", "r2", "r3"), times = 4)
```

```{r}
pcaData <- plotPCA(
  vsd,
  intgroup = c("design"),
  returnData = TRUE
)

percentVar <- round(100 * attr(pcaData, "percentVar"))

ggplot(
  pcaData,
  aes(
    x = PC1,
    y = PC2,
    colour = design
  )
) +
  geom_point(size = 3) +
  theme_bw() +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  labs(title = "A")
```
```{r}
metadata$replicate <- rep(c("r1", "r2", "r3"), times = 4)
```


```{r}
pcaData <- plotPCA(
  vsd,
  intgroup = c("design", "replicate"),
  returnData = TRUE
)

ggplot(
  pcaData,
  aes(
    x = PC1,
    y = PC2,
    colour = design,
    shape = replicate
  )
) +
  geom_point(size = 3) +
  theme_bw() +
  xlab(paste0("PC1: ", percentVar[1], "% variance")) +
  ylab(paste0("PC2: ", percentVar[2], "% variance")) +
  labs(title = "A")
```

```{r}
sampleDists <- dist(
  t(assay(vsd)),
  method = "euclidean"
)

sampleDistMatrix <- as.matrix(sampleDists)

rownames(sampleDistMatrix) <- colnames(vsd)
colnames(sampleDistMatrix) <- colnames(vsd)

pheatmap(
  sampleDistMatrix,
  clustering_method = "complete",
  main = "B"
)
```

